---
title: "Example: Taxonomic, geographic, and temporal analyses of fungal fire association"
author: "Hunter J. Simpson"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
## Introduction

The following R code is based on the methods used by Simpson and Schilling (2021).

## Download MyCoPortal data

Data is first retrieved from the MyCoPortal (https://mycoportal.org/portal/index.php) using either `mycoportal_tab` or by manually downloading data using the MyCoPortal web interface. If downloaded manually, datasets can then be imported into R using `read.delim`.

```{r data_download, include=TRUE, eval=FALSE, echo=TRUE}
#Load package
library(fungarium) 

#Download all US records for fungi in the subphylum Agaricomycotina
agaricomycotina <- mycoportal_tab("/path/to/directory", "Agaricomycotina", 
                                  taxon_type = "4", country = "United States") 

#Download all US records for fungi in the subphylum Pezizomycotina
pezizomycotina <- mycoportal_tab("/path/to/directory", "Pezizomycotina", 
                                 taxon_type = "4", country = "United States") 

#Combine datasets
all_fungi <- rbind(agaricomycotina, pezizomycotina) 
```

## Update taxon names and clean up

Some MyCoPortal records will have outdated or invalid taxonomy; therefore, all taxon names need to be either updated or confirmed using currently accepted taxonomic classification.

After updating taxonomy (in the example below), some records will now be classified outside of Agaricomycotina and Pezizomycotina. These records need to be removed. The filtering step shown below would be easier if subphyla information was listed in the output of taxon_update, but the GBIF utility used by taxon_update only outputs primary taxonomic ranks; therefore, subphyla cannot be used directly for filtering.

Note that taxon name updating and filtering was done prior to the taxonomic, geographic, and temporal analyses done by Simpson and Schilling (2021). This is not necessarily required for geographic and temporal analyses; however, you run the risk of including "bad taxa" (i.e. taxa that are not actually classified within your originally specificied taxonomic group) in your analyses. Thus it is recommended to update taxon names and filter before performing any downstream analyses.

```{r taxon_name_update, include=TRUE, eval=FALSE, echo=TRUE}
#Update names and classification
all_fungi <- taxon_update(all_fungi, species_only=T) 

#Remove records not in the phyla Ascomycota and Basidiomycota
all_fungi <- all_fungi[all_fungi$new_phylum %in% c("Ascomycota", "Basidiomycota"),] 

#Remove records that are classified in classes not within Agaricomycotina and Pezizomycoyina
all_fungi <- all_fungi[!all_fungi$new_class %in% c("Ustilaginomycetes", 
                                                   "Pucciniomycetes",
                                                   "Agaricostilbomycetes",
                                                   "Atractiellomycetes",
                                                   "Cystobasidiomycetes",
                                                   "Microbotryomycetes",
                                                   "Exobasidiomycetes",
                                                   "Neolectomycetes",
                                                   "Saccharomycetes", 
                                                   "Wallemiomycetes", 
                                                   ""),] 

#Remove records with taxon names that could not be updated or confirmed
all_fungi <- all_fungi[all_fungi$new_name!="",]
```

## Find fire-associated records

Fire-associated records are identified using a complex regex search string of fire-related keywords and phrases. The positive search string helps find fire-relevant records, while the negative string helps remove records that were falsely identified as fire associated. The negative string is not always necessary, escpecially if the positive string has been thoroughly optimized.

```{r find_trait, include=TRUE, eval=FALSE, echo=TRUE}
#Specify "positive" search string
string1 <- "(?i)charred|(?i)burn(t|ed)|(?i)scorched|(?i)fire.?(killed|damaged|scarred)|(?i)killed.by.fire"

#Specify "negative" search string
string2 <- "(?i)un.?burn(t|ed)"

#Find fire-associated records
fire_fungi <- find_trait(all_fungi, pos_string=string1, neg_string=string2)
```

## Calculate taxon-specific fire-associated enrichment factors

To compare the extent of fire association between taxa, fire-associated enrichment factors are calculated by dividing the number of fire-associated records by the number of total records for each taxon. Larger enrichment values indicate stronger fire association (i.e. a higher proportion of fire-associated records) for that taxon, in comparison to taxa with smaller values.

```{r taxon_enrich, include=TRUE, eval=FALSE, echo=TRUE}
#Count the frequency of each taxon in the total dataset
counts1 <- plyr::count(all_fungi,vars=c("new_name", "new_author", 
                                       "new_full_name", "new_kingdom",
                                       "new_phylum", "new_class", "new_order",
                                       "new_family", "new_genus",
                                       "new_specific_epithet"))

#Count the frequency of each taxon in the fire-associated records dataset
counts2 <- plyr::count(fire_fungi,vars=c("new_full_name"))
colnames(counts2)[2] <- "fire_freq"

#Merge total frequency and fire frequency data into one dataframe
counts <- merge(counts1, counts2, by.x="new_full_name", 
                by.y="new_full_name", all.x=T, all.y=T)
counts[is.na(counts$fire_freq),"fire_freq"] <- 0

#Use total freq and fire freq to calculate enrichment factors
counts$fire_ratio <- counts$fire_freq/counts$freq

#(Optional) to increase confidence in calculated enrichment factors, a filtering threshold can be used to remove taxa that have not been thoroughly sampled
counts <- counts[counts$freq>=5,]#removes taxa that have less than 5 total records
```

## Calculate county-specific fire-associated enrichment factors

To compare the extent of fungal fire association between counties, fire-associated enrichment factors are calculated by dividing the number of fire-associated records by the number of total records for each county. Larger enrichment values indicate stronger fire association (i.e. a higher proportion of fire-associated records) for that county, in comparison to counties with smaller values.

```{r county_enrich, include=TRUE, eval=FALSE, echo=TRUE}
#Fix mispelled county names and assign fips codes
all_fungi <- get_fips(all_fungi, assign_counties=T)

#Count the frequency of counties in total dataset
counts1 <- plyr::count(all_fungi,vars=c("county_fips"))

#Count the frequency of counties in fire dataset
counts2 <- plyr::count(fire_fungi,vars=c("county_fips"))
colnames(counts2)[2] <- "fire_freq"

#Merge total frequency and fire frequency data into one dataframe
counts <- merge(counts1, counts2, by.x="county_fips", 
                by.y="county_fips", all.x=T, all.y=T)
counts[is.na(counts$fire_freq),"fire_freq"] <- 0

#Use total freq and fire freq to calculate enrichment factors
counts$fire_ratio <- counts$fire_freq/counts$freq

#(Optional) to increase confidence in calculated enrichment factors, a filtering threshold can be used to remove counties that have not been thoroughly sampled
counts <- counts[counts$freq>=5,]#removes counties that have less than 5 total records
```

## Calculate year-specific fire-associated enrichment factors

To examine changes in fungal fire association over time, fire-associated enrichment factors are calculated by dividing the number of fire-associated records by the number of total records for each year. Larger enrichment values indicate stronger fire association (i.e. a higher proportion of fire-associated records) for that year, in comparison to years with smaller values.

```{r year_enrich, include=TRUE, eval=FALSE, echo=TRUE}
#Filter out records with no year or not within a certain year range
all_fungi <- all_fungi[all_fungi$year!="",] #remove blanks
all_fungi$year_int <- as.integer(all_fungi$year) #convert years to integers
all_fungi <- all_fungi[all_fungi$year_int %in% c(1800:2019),] #filter out records not in the year range of 1800 to 2019

#Count the frequency of each year in total dataset
counts1 <- plyr::count(all_fungi,vars=c("year"))

#Count the frequency of each year in fire dataset
counts2 <- plyr::count(fire_fungi,vars=c("year"))
colnames(counts2)[2] <- "fire_freq"

#Merge total frequency and fire frequency data into one dataframe
counts <- merge(counts1, counts2, by.x="year", 
                by.y="year", all.x=T, all.y=T)
counts[is.na(counts$fire_freq),"fire_freq"] <- 0

#Use total freq and fire freq to calculate enrichment factors
counts$fire_ratio <- counts$fire_freq/counts$freq

#(Optional) to increase confidence in calculated enrichment factors, a filtering threshold can be used to remove years that have not been thoroughly sampled
counts <- counts[counts$freq>=5,]#removes years that have less than 5 total records; likely more useful for shorter time intervals, as most years are likely to have many records
```

## References

Simpson, HJ, Schilling, JS. 2021. Using aggregated field collections data and the novel R package FUNGARIUM to investigate fungal fire association. Mycologia. (IN REVIEW)
