---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
fungarium
======

<!-- badges: start -->
![GitHub](https://img.shields.io/github/license/hjsimpso/fungarium)
![Github commits](https://img.shields.io/github/commits-since/hjsimpso/fungarium/latest)
[![Maintenance](https://img.shields.io/badge/Maintained%3F-yes-green.svg)](https://GitHub.com/hjsimpso/fungarium/graphs/commit-activity)

<!-- badges: end -->

An R package for enhancing the utility of fungarium collections and community science observations in ecological analyses of fungi. 

## Installing fungarium
```{r installation, include=TRUE, eval=FALSE, echo=TRUE}
install.packages("remotes") #install 'remotes' (if not already installed)
remotes::install_github("hjsimpso/fungarium@*release") #install the latest fungarium release
#to install the latest development version use:
remotes::install_github("hjsimpso/fungarium@main")
```

## Meta
* Please cite: Simpson, H., Schilling, J. 2021. Using aggregated field collection data and the novel R package FUNGARIUM to investigate fungal fire-association. Mycologia 113(4):842-855. DOI: 10.1080/00275514.2021.1884816
* This package is released with a [Contributor Code of Conduct](https://github.com/hjsimpso/fungarium/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
* Please report any bugs.
