---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
fungarium
======

<!-- badges: start -->
![GitHub](https://img.shields.io/github/license/hjsimpso/fungarium)
[![GitHub release](https://img.shields.io/github/release/hjsimpso/fungarium.svg)](https://GitHub.com/hjsimpso/fungarium/releases/)
![Github commits](https://img.shields.io/github/commits-since/hjsimpso/fungarium/latest)
[![Maintenance](https://img.shields.io/badge/Maintained%3F-yes-green.svg)](https://GitHub.com/hjsimpso/fungarium/graphs/commit-activity)

<!-- badges: end -->

An R package for enabling the use of fungaria collections data in comprehensive trait analyses. 

## Installing fungarium
```{r installation, include=TRUE, eval=FALSE, echo=TRUE}
install.packages("devtools") #Install 'devtools' (if not already installed)
devtools::install_github("hjsimpso/fungarium") #Retrieve fungarium package from github repository and install
```

## Installing Docker
[Docker](https://docs.docker.com/get-started/overview/) needs to be installed on your system in order to utilize the `mycoportal_tab` function. All other functions in the **fungarium** package do not require Docker, so if you do not wish to utilize `mycoportal_tab` you do not need to install Docker.


Instructions for installing Docker (available for Linux platforms, macOS, and Windows 10): https://docs.docker.com/engine/install/

## Examples and detailed guidance on package usage
See [fungarium book](https://hjsimpso.github.io/fungarium_book/).

## Meta
* Get citation information for fungarium in R via `citation(package = 'fungarium')`
* This package is released with a [Contributor Code of Conduct](https://github.com/hjsimpso/fungarium/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
* Please report any bugs.
