---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# fungarium

<!-- badges: start -->
<!-- badges: end -->

An R package for enhancing the access and analysis of fungaria collections data. 

## Introduction

### Installing fungarium

```{r installation, include=TRUE, eval=FALSE, echo=TRUE}
install.packages("devtools") #Install 'devtools' (if not already installed)
devtools::install_github("hjsimpso/fungarium") #Retrieve fungarium package from github repository and install
```

### Installing Docker

Docker software (https://docs.docker.com/get-started/overview/) needs to be installed on your system in order to utilize the `mycoportal_tab` function. All other functions in the fungarium package do not require Docker, so if you do not wish to utilize `mycoportal_tab` you do not need to install Docker.


Instructions for installing Docker (available for Linux platforms, macOS, and Windows 10): https://docs.docker.com/engine/install/

## Retrieving MyCoPortal datasets

Manually interacting with the MyCoPortal web interface (https://mycoportal.org/portal/index.php) can be a time consuming process, especially if one needs to submit a large variety of separate queries for various taxa, locations, dates, etc. The `mycoportal_tab` function provides a convenient alternative by allowing users to submit queries and download complete datasets from entirely within the R programming environment.

Examples:
```{r mycoportal_tab, include=TRUE, eval=FALSE, echo=TRUE}
# Check that Docker is installed and running on your system

# Load fungarium
library(fungarium) 

# Submit queries to MyCoPortal and retrieve complete datasets
# MyCoPortal tab-delmited file is saved in specified directory and then imported into R as a data.frame
query1 <- mycoportal_tab("/path/to/directory", "Agaricus", taxon_type = "1", country = "United States", state="Minnesota", messages = F, rec_numb = F) 
query1[1:10, c("scientificName", "country", "stateProvince")] # Preview dataset

query2 <- mycoportal_tab("/path/to/directory", "Russulaceae", taxon_type = "2", country = "United States", state="Minnesota", messages = F, rec_numb = F) 
query2[1:10, c("scientificName", "country", "stateProvince")] # Preview dataset

query3 <- mycoportal_tab("/path/to/directory", "Polyporales", taxon_type = "4", country = "United States", state="Minnesota", messages = F, rec_numb = F) 
query3[1:10, c("scientificName", "country", "stateProvince")] # Preview dataset

```

## Meta
* Get citation information for fungarium in R via `citation(package = 'taxize')`
* This package is released with a [Contributor Code of Conduct](https://github.com/hjsimpso/fungarium/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
